---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata/data/activity.zip"
download.file(fileUrl, destfile="./RepResActivity.zip")
unzip("./RepResActivity.zip")
activity <- read.csv("activity.csv")
activity$date <- as.Date(activity$date, "%Y-%m-%d")
# looking at the structure:
str(activity)
```


## What is mean total number of steps taken per day?
```{r}
# aggregating steps per day for mean and median calculation
dailysteps <- aggregate(activity$steps, by=list(Date = activity$date), sum)
colnames(dailysteps)[2] <- "sumofdailysteps"

hist(dailysteps$sumofdailysteps, main = "histogram of sum of number of daily steps")
mean(dailysteps$sumofdailysteps, na.rm = TRUE) # mean of daily steps
median(dailysteps$sumofdailysteps, na.rm = TRUE) # median of daily steps
```


## What is the average daily activity pattern?
```{r}
meanintervsteps <- aggregate(activity$steps, by=list(interval = activity$interval), mean, na.rm = TRUE)
colnames(meanintervsteps)[2] <- "meanofintervsteps"
plot(x=meanintervsteps$interval, y=meanintervsteps$meanofintervsteps, type = 'l')
max(meanintervsteps)
which.max(meanintervsteps$meanofintervsteps)
meanintervsteps[104,]
```
The maximum is inside row 104, so it's the interval 835 -> looking at the plot this seems to be correct


## Imputing missing values
```{r}
impactivity <- activity # new data set which will get imputed data
sum(is.na(impactivity$steps)) # checking how many NAs there are right now
for(i in 1:(nrow(impactivity))) { # replacing each NA with the mean of the corresponding interval
    if(is.na(impactivity$steps[i])) {
        impactivity$steps[i] = meanintervsteps$meanofintervsteps[meanintervsteps$interval==impactivity$interval[i]]
    }
}
# check if an na is left
sum(is.na(impactivity$steps))

#### Recalculating the daily steps, the histogram and mean and median for the data set with imputed data
impdailysteps <- aggregate(impactivity$steps, by=list(Date = impactivity$date), sum)
colnames(impdailysteps)[2] <- "sumofdailysteps"

hist(impdailysteps$sumofdailysteps, main = "histogram of sum of number of daily steps")

mean(dailysteps$sumofdailysteps, na.rm = TRUE) # before imputing
median(dailysteps$sumofdailysteps, na.rm = TRUE) # before imputing
mean(impdailysteps$sumofdailysteps, na.rm = TRUE) # after imputing
median(impdailysteps$sumofdailysteps, na.rm = TRUE) # after imputing
```
The means stay the same - median changes a bit and is now equal to mean


## Are there differences in activity patterns between weekdays and weekends?
```{r}
impactivity$daytype <- weekdays(impactivity$date)
for(i in 1:nrow(impactivity)) {
    if(impactivity$daytype[i] == "Samstag" || impactivity$daytype[i] == "Sonntag") {
        impactivity$daytype[i] <- "weekend"
    } else {
        impactivity$daytype[i] <- "weekday"
    }
}
impactivity$daytype <- as.factor(impactivity$daytype)

impmeanintervsteps <- aggregate(steps ~ interval + daytype, impactivity, mean)
library(ggplot2)
qplot(x = interval, y = steps, data = impmeanintervsteps, geom = ("line")) + facet_grid(~ daytype)
```